---
title: "MAP535 Project"
author: "Leonardo Natale - Guillaume Le Fur"
date: "08/12/2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(dplyr)
library(plotly)
library(leaps)
library(MASS)
knitr::opts_chunk$set(echo = FALSE)

train <- as.data.frame(data.table::fread("data/train_preprocessed.csv"))
numeric_train <- data.frame(train %>% select_if(is.numeric))
```

# Introduction



The following table contains the 5 predictors that are the most correlated to the sales price.



# Exploratory Data Analysis/Inital Modeling

```{r High Correlations}
cn <- colnames(numeric_train)
df.high.cor <- data.frame(
  x = c(),
  y = c(),
  cor = c(),
  stringsAsFactors = F
)
cor.threshold <- .7
for (i in 1:ncol(numeric_train)){
  cni <- cn[i]
  for (j in 1:ncol(numeric_train)){
    cnj <- cn[j]
    cv <- cor(numeric_train[[cni]], numeric_train[[cnj]])
    if (cv > cor.threshold && i < j){
      df.high.cor <- df.high.cor %>% rbind(
        data.frame(
          x = cni, 
          y = cnj, 
          cor = cv
        )
      )
    }
  }
}
df.high.cor <- df.high.cor %>% arrange(desc(cor))
df.high.cor
```

```{r Correlations with Y}
apply.res <- apply(numeric_train, 2, function(x){cor(x, numeric_train$SalePrice)})
df.cor.to.y <- data.frame(
  column = names(apply.res),
  correlation = apply.res
) %>% 
  filter(correlation < 1) %>% 
  top_n(5, correlation)
df.cor.to.y

```

# Modeling and Diagnostics

```{r Full LM}
res.lm.intercept <- lm(SalePrice~1, data = train)
res.full.lm <- lm(SalePrice~., data = train)
sum.full.lm <- summary(res.full.lm)
nrow(sum.full.lm$coefficients)
```

The largest p-value for the T-tests of every predictor is `r min(sum.full.lm$coefficients[, 4])`, which is really high. On the model with all the data, the significativity of the predictors isn't really high.

```{r Variable selection}
cn <- colnames(train)
cn <- cn[cn != "SalePrice"]
res.backward <- stepAIC(
  res.full.lm, 
  ~.,
  trace = F,
  direction=c('backward')
)
nrow(summary(res.backward)$coefficients)
```


# Final Models

# Discussion

